### 2018-07-03 submission/manuscript.Rmd
change title
update abstract - background and methods

# Error Introduced into 16S rRNA Gene Sequencing Varies by High Fidelity DNA Polymerase Used

@@

**Background.** It is challenging to compare 16S rRNA gene sequencing data across studies and one of the reasons for this is due to error. There are many different places throughout the workflow where error can be introduced into the pipeline. Here, we focus on studying how the number of cycles and high fidelity (HiFi) DNA polymerase introduce error by varying cycle number and polymerase used to amplify 16S rRNA genes in human fecal and mock community samples.

**Methods.** We extracted DNA from fecal samples (n=4) using a PowerMag DNA extraction kit with a 10 minute bead beating step and amplified at 15, 20, 25, 30, and 35 cycles using Accuprime, Kappa, Phusion, Platinum, or Q5 HiFi DNA polymerase. Amplification of mock communities (technical replicates n=4) consisting of previously isolated whole genomes of 8 different bacteria used the same approach. The analysis initially examined the number of Operational Taxonomic Units (OTUs) for fecal samples and mock communities. It also assessed polymerase and number of cycle dependent differences in the Bray-Curtis index, error rate, sequence error prevalence, chimera prevalence, and the  correlation between chimera prevalence and number of OTUs.

>>>>

# Assessing Differences in High Fidelity DNA Polymerase for 16S rRNA Gene Sequencing

@@

**Background.** A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, amplification, and sequencing steps. At each of these stages error can be introduced that will change the underlying bacterial community composition results. In this study we focus on the amplification step's contribution to this overall error. To accomplish this we assessed 16S rRNA gene sequencing results in human fecal and mock community samples after using different high fidelity (HiFi) DNA polymerases and number of amplification cycles.

**Methods.** We extracted DNA from fecal samples (n=4) using a PowerMag DNA extraction kit with a 10 minute bead beating step and amplified at 15, 20, 25, 30, and 35 cycles using Accuprime, Kappa, Phusion, Platinum, or Q5 HiFi DNA polymerase. Amplification of mock communities (technical replicates n=4) consisting of previously isolated whole genomes of 8 different bacteria used the same approach. We first assessed error rate, sequence error prevalence, chimera prevalence, correlation between chimera prevalence and number of Operational Taxonomic Units (OTUs), and GC dependent differences by polymerase and number of cycles. Next, differences in the number of OTUs and taxa was examined based on the polymerase and number of cycles used. Additionally, differences in the bacterial community composition by the Bray-Curtis index also was assessed based on polymerase and number of cycles. Finally, Random Forest models were created to test whether the bacterial community was better at classifying polymerases, number of cycles, or individual donor. We also assessed whether the most important taxa in the polymerase and number of cycle Random Forest models also were the most important in the model for individual donors.
